---
title: "Scraping Data"
author: "Yusuf Ameri"
date: "December 3, 2016"
output: html_document
---

### Load Neccessary Libraries
```{r libraries, error=FALSE, warning=FALSE,message=FALSE,results='hide'}
library(rvest)    # for web scraping
library(dplyr)    # for column selection
library(tidyr)    # for column tidying
library(magrittr) # for piping
library(tibble)   # for pretty table printing
library(readr)    # for converting column types
```


### Scrape url from Wiki
We need to scrap an html table from wiki that has info on presidential elections and popular votes. This table only has popular vote data from 1824. Click [Here](https://en.wikipedia.org/wiki/List_of_United_States_presidential_elections_by_popular_vote_margin#List) To see the wiki yourself.
```{r main}
# URL of wiki of list of presidents
url <- "https://en.wikipedia.org/wiki/List_of_United_States_presidential_elections_by_popular_vote_margin"

# read URL
president_list <- read_html(url)

# Scrape following the xpath for the table
president_table <- president_list %>% html_node(xpath = '//*[@id="mw-content-text"]/table[2]') %>% html_table()

# display first 2 rows and first 4 attributes
head(president_table[1:2,1:4])
```

As you can see above, the data table is not yet clean and has some work to do, we will first remove the first two rows which are simply headers and an empty line.
```{r clean1}
# remove the first 2 rows which are headers
president_table <- president_table[3:nrow(president_table),]

# display first 2 rows and first 4 attributes
head(president_table[1:2,1:4])
```

Next, we will rename the columns so that they make sense.
```{r rename_columns}
# rename columns
colnames(president_table)[1] <- "Election_Number"
colnames(president_table)[2] <- "Election_Year"
colnames(president_table)[3] <- "Winner"
colnames(president_table)[4] <- "W_Party"
colnames(president_table)[5] <- "Electoral_Proportion"
colnames(president_table)[6] <- "Electoral_Percentage"
colnames(president_table)[7] <- "Popular_Percentage"
colnames(president_table)[8] <- "Popular_Margin"
colnames(president_table)[9] <- "Winner_Votes"
colnames(president_table)[10] <- "Winner_Votes_Margin"
colnames(president_table)[11] <- "Runner-Up"
colnames(president_table)[12] <- "L_Party"
colnames(president_table)[13] <- "Turnout_pct"

# store president_table as a tibble for better printing
president_table <- tbl_df(president_table)

# display president table with new column names
head(president_table)
```
Things are starting to look much better. We still need to change the the datatypes of many of this columns and extract more useful attributes given what we have (for example, number of votes for losing opponent)

```{r pres_str}
# lets have a look at the datatypes in each columns
str(president_table)
```

We need to convert Some of these column types to the right datatype
```{r convert_cols}
# convert column types
president_table <- president_table %>% readr::type_convert(col_types = cols(
  Election_Number = col_integer(),
  W_Party = col_factor(levels(president_table$W_Party %>% as.factor())),
  Winner = col_factor(levels(president_table$Winner %>% as.factor())),
  `Runner-Up` = col_factor(levels(president_table$`Runner-Up` %>% as.factor())),
  L_Party = col_factor(levels(president_table$L_Party %>% as.factor()))
))
```

For columns with string representations of integers and percentage points, we have to strip unneccesary characters such as the percent '%' sign and division sign '/'.
Here are some examples
```{r examples}
# need to get rid of / sign
president_table$Electoral_Proportion[1]

# need to get rid of % sign
president_table$Turnout_pct[1]
```

In the process of getting rid of the 'Electoral_Proportion' Column, we will create new columns for Electoral votes won by the winning candidate and total electoral votes in the election year. It will be up to the user to query what percentage the electoral vote the winner recieved (**for now**).

```{r new_columns}
# replace electoral proportion into electoral votes won (W_Electoral_Votes) and Total_Electoral_Votes
president_table <- 
  president_table %>% 
  tidyr::separate(
    col = "Electoral_Proportion", 
    into = c("W_Electoral_Votes", "Total_Electoral_Votes"), 
    sep = "/")

# convert those columns into integers
president_table <- president_table %>% readr::type_convert(col_types = cols(
  W_Electoral_Votes = col_integer(),
  Total_Electoral_Votes = col_integer()
))

# remove the electoral percentage column
president_table <- 
  president_table %>% 
    select(-Electoral_Percentage)

# take a look
president_table %>% select(Winner, W_Electoral_Votes, Total_Electoral_Votes) %>% head()
```

### Convert Voter Turnout Percentage to a Double
```{r voter_turnout}
# remove % sign and special characters
president_table <- 
  president_table %>% 
    tidyr::separate(col="Turnout_pct",
                    into = "Turnout_pct",
                    sep = "%",
                    remove = TRUE,
                    extra = "drop")

# convert voter turnout into a double and divide by 100
president_table$Turnout_pct <- president_table$Turnout_pct %>% as.double()
president_table <- president_table %>% mutate(Turnout_pct = Turnout_pct/100)
```

### Account for Losers Vote Count
Instead of having a column that displays the margin of the winners victory, we would like to rather have a column that displays how many votes the losing candidate recieved. But first, however, we need to convert the column of Winner_Votes into an integer. This involves cleaning (removing commas and other characters) from each row.
```{r loser_votes}
# First remove commas and random characters from Winner_Votes Column
president_table$Winner_Votes <- gsub("(\\W)","", president_table$Winner_Votes)

# Next, convert to integer
# president_table$Winner_Votes %>% as.integer()
```

# ABORT, DOUG ALREAD DID ABOVE :(

### Scrape Population/Census Table
```{r scrape_census}
url <- 'https://en.wikipedia.org/wiki/Demographic_history_of_the_United_States#Historical_population'
xpath <- '//*[@id="mw-content-text"]/table[2]'

census <- read_html(url)

census_table <- census %>% html_node(xpath = xpath) %>% html_table()

# remove last row
census_table <- census_table[2:nrow(census_table),]

# pdf source: http://www2.census.gov/prod2/statcomp/documents/CT1970p2-13.pdf
```

### Stip commas from numbers
```{r strip_commas}
census_table$Population <- gsub(",","", census_table$Population)
```



Now lets write the table as an SQL entry file
```{r write_to_sql}
# Now let's write the census to a file as a sql command
fileConn <- "population.txt"
write("INSERT INTO `population` (`year`, `population`) VALUES ", fileConn)
for (k in 1:dim(census_table)[1]){
  if (k == length(partyList)) {
    write(sprintf("('%s', '%s'), ", census_table[k,1],census_table[k,2]), fileConn, append=TRUE)
  } else {
    write(sprintf("('%s', '%s')", census_table[k,1],census_table[k,2]), fileConn, append=TRUE)
  }
}
```

